---
title: "Working with USGS Data and Functions"
author: "Juli Scamardo"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(dataRetrieval)
library(dygraphs) #Time series plotting
library(tidyr) 
library(xts) #Time series objects
library(ggthemes)

knitr::opts_chunk$set(echo = TRUE)
```

# Working with USGS data

## Download USGS data using dataRetrieval


#### Non function oriented method

```{r,eval=F}
lincoln_q <- readNWISdv('06752260', parameterCd = '00060',
                        startDate = '1975-10-01',endDate = '2019-10-01') %>%
  rename(q_cfs = X_00060_00003,
         q_cd = X_00060_00003_cd) %>%
  mutate(site = 'lincoln')

#summary(lincoln_q)

elc_q <- readNWISdv('06752280', parameterCd = '00060',
                        startDate = '1975-10-01',endDate = '2019-10-01') %>%
  rename(q_cfs = X_00060_00003,
         q_cd = X_00060_00003_cd) %>%
  mutate(site = 'elc')


combo_data <- bind_rows(lincoln_q,elc_q)
summary(combo_data)
#summary(elc_q)
```


#### Function oriented method


```{r}

q_downloader <- function(site_no = '06752260',
                         site_id = 'lincoln'){
  df <- readNWISdv(site_no, 
                   parameterCd = '00060',
                   startDate = '1950-10-01',
                   endDate = '2019-10-01') %>%
  rename(q_cfs = X_00060_00003,
         q_cd = X_00060_00003_cd) %>%
    mutate(site = site_id)
  return(df)
}


lincoln_q <- q_downloader(site_no = '06752260',
                          site_id = 'lincoln')
elc_q <- q_downloader(site_no = '06752280',
                      site_id = 'elc')
#
combo_data <- bind_rows(lincoln_q,elc_q)

summary(combo_data)
```


## Plot the discharge data

#### Time series (xaxis = date)

```{r}

ggplot(combo_data,aes(x=Date,y=q_cfs,color=site)) + 
  geom_line() 


```

#### Density plot 

```{r}
ggplot(combo_data,aes(x=q_cfs,color=site)) +
  geom_density() +
  scale_x_log10()

```


#### Interactive graphs 

```{r,eval=F}
#Structore examination
#str(combo_data)
wide_data <- combo_data %>%
  select(-agency_cd,-site_no,-q_cd) %>%
  spread(key=site,value=q_cfs)

wide_xts <- xts(wide_data %>%
                  select(elc,lincoln),
                order.by = wide_data$Date)

dygraph(wide_xts)
```


#### Function for interactive graph

```{r}

xts_maker <- function(df){
  wide_data <- df %>%
    select(-agency_cd,-site_no,-q_cd) %>%
    spread(key=site,value = q_cfs)

  wide_xts <- xts(wide_data %>%
                  select(-Date),
                order.by = wide_data$Date)
  
  return(wide_xts)
  
}


wide_xts <- xts_maker(combo_data)
wide_elc <- xts_maker(elc_q)


dygraph(wide_xts)
dygraph(wide_elc)
```


# In class work

## Write a function that makes ggplot time series graph of discharge

Here I want you to write your own function that adds custom 
themes to your normal ggplot call (so you don't have to write
out theme_few and theme(legend.position = ...)) over and over again. You may need to load the library `ggthemes`. 

This function call should generate a clean time series plot 
with a single command and a single option (which data frame to 
plot). Please plot the `combo_data` dataset here.


```{r}
## Create function that plots time series in ggplot2
plot_ts <- function(df){
  p <- ggplot(data = df, aes(x = Date, y = q_cfs, color = site))+ 
    geom_line() + 
    theme_few() + 
    scale_color_economist() + 
    labs(x = 'Date', y = 'Discharge (cfs)') + 
    theme(legend.position = c(0.1, 0.85))
  
  return(p)
}

## plot combo data using function above
plot_ts(combo_data)

```


## Download discharge data
Here, I downloaded data for the Rio Grande River through Big Bend National Park. I'm looking at 2 USGS gages: one upstream at Castolon and one downstream at Rio Grande Village. These two gages bracket streamflow through Big Bend. 

The 'q_downloader' function worked for this download, but __'q_downloader' could be brittle because certain variable names and values are hard coded. Some sites many not follow this naming scheme.__

```{r}
## Downloading data for Rio Grande near Castolon Visitors Center
castolon <- q_downloader(site_no = '08374550', 
                         site_id = 'castolon')

## Downloading data for Rio Grande near Rio Grande Village
rgvillage <- q_downloader(site_no = '08375300', 
                          site_id = 'rgvillage')

```


## Joining your two datasets
My goal was to join the two datasets in a way that results in a wide dataset that can be converted to a time series and interactive graph. 

```{r}
## function for joining discharge datasets resulting in a wide dataset
q_join <- function(df1, df2) {
  
  # gather site name of both sites
  name1<- sym(df1$site[nrow(df1)])
  name2<- sym(df2$site[nrow(df2)])
  
  # join datasets and rename columns
  q <- full_join(df1, df2, by = 'Date') %>%
    select(agency_cd.x, site_no.x, site_no.y, Date, q_cfs.x, q_cfs.y) %>%
    rename(agency_cd = 1, site_no_1 = 2, site_no_2 = 3, date = 4, 
           !!name1 := 5, !!name2 :=6)

return(q)
}

## Creating a wide dataset for streamflow through Big Bend NP
bbnp <- q_join(df1 = castolon, df2 = rgvillage)

## Looking at the top of the dataset to see if it joined correctly
head(bbnp)
```


## Plotting the data
The dataset created above can then be plotted as an interactive graph using dygraphs. Here, I created a function that formats the data into a time series and then plots it. 


```{r}
## Function for formatting and plotting discharge time series
dy_builder <- function(df, title){
  # slim down df
  w <- df %>%
    select(-agency_cd,-contains("site_no")) 
  
  # create time series
  w_xts <- xts(w %>% select(-date), order.by = w$date)
  
  # plot on interactive map
  g <- dygraph(w_xts, main = title, ylab = 'discharge (cfs)') %>%
    dyOptions(colors = colorspace::diverge_hcl(2))
  
  return(g)
}

## Plot the bbnp data using the function
dy_builder(bbnp, 'Rio Grande Streamflow: Big Bend NP')

```


